---
title: "Skye Vegetation Survey: Data Visualisation"
author: "Bhagyesh Sagole"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)

# Find the project from this Rmd file, whether chunks are run interactively
# in RStudio or the whole document is knitted.
rmd_path <- if (interactive() && requireNamespace("rstudioapi", quietly = TRUE) &&
                rstudioapi::isAvailable()) {
  rstudioapi::getSourceEditorContext()$path
} else {
  knitr::current_input(dir = TRUE)
}

project_root <- if (nzchar(rmd_path)) {
  dirname(dirname(normalizePath(rmd_path)))
} else {
  normalizePath("..")
}
```

# Read in the data

The main dataset contains one row for each species-by-releve observation.

```{r read-data}
survey_raw <- read_csv(
  file.path(project_root, "output", "output.csv"),
  col_types = cols(
    releve_id = col_character(),
    latitude = col_double(),
    longitude = col_double(),
    altitude_ft = col_double(),
    altitude_m = col_double(),
    domin_value = col_double(),
    domin_cover_min_pct = col_double(),
    domin_cover_max_pct = col_double(),
    presence_binary = col_logical(),
    needs_review = col_logical(),
    .default = col_character()
  )
)
```

# First look at the data

```{r preview}
head(survey_raw)
```

```{r dimensions}
dim(survey_raw)
```

```{r structure}
glimpse(survey_raw)
```

# Basic data checks

## Missing values by column

```{r missing-values}
missing_values <- survey_raw %>%
  summarise(across(everything(), ~ sum(is.na(.x)))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "n_missing"
  ) %>%
  arrange(desc(n_missing))

missing_values
```

## Dataset summary

```{r basic-summary}
survey_raw %>%
  summarise(
    observation_rows = n(),
    tables = n_distinct(table_id),
    releves = n_distinct(paste(table_id, releve_id)),
    species = n_distinct(species, na.rm = TRUE),
    classes = n_distinct(class, na.rm = TRUE),
    associations = n_distinct(association, na.rm = TRUE),
    geolocated_rows = sum(!is.na(latitude) & !is.na(longitude))
  )
```

## Numeric variables

```{r numeric-summary}
survey_raw %>%
  select(latitude, longitude, altitude_ft, altitude_m,
         domin_value, domin_cover_min_pct, domin_cover_max_pct) %>%
  summary()
```

# Next steps

The next stage can clean and reshape the data before creating visualisations of
species frequency, vegetation classes, altitude, and geographic distribution.
