
---
title: "Isle of Skye Vegetation Survey — Birks (1973)"
subtitle: "Exploratory Data Analysis"
author: "Bhagyesh Sagole & Pawan Kumar"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)
```
```{r libraries, message=FALSE}
library(tidyverse)
library(readr)
library(knitr)
library(kableExtra)
library(vegan)
library(DT)
library(plotly)
library(RColorBrewer)
library(indicspecies)
library(permute)
library(patchwork)
library(glue)
library(scales)

theme_skye <- function(base_size = 11) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      panel.grid.minor  = element_blank(),
      panel.grid.major  = element_line(colour = "#eeeeee"),
      axis.text         = element_text(colour = "#444444"),
      plot.title        = element_text(face = "bold", size = base_size + 1,
                                       colour = "#2e5e4e", margin = margin(b = 6)),
      plot.subtitle     = element_text(size = base_size - 1, colour = "#666666",
                                       margin = margin(b = 10)),
      legend.text       = element_text(size = base_size - 2),
      strip.text        = element_text(face = "bold", size = base_size - 1)
    )
}
```

---

# 1 · Survey Overview

This section loads the data, applies cleaning steps agreed with the PI,
and reports top-level dataset metrics.

```{r load-data}
# ── Load raw data ─────────────────────────────────────────────────────────────
survey_raw <- read_csv(
  "~/Desktop/Gavin briks lab/output.csv",
  col_types = cols(
    releve_id            = col_character(),
    latitude             = col_double(),
    longitude            = col_double(),
    altitude_ft          = col_double(),
    altitude_m           = col_double(),
    domin_value          = col_double(),
    domin_cover_min_pct  = col_double(),
    domin_cover_max_pct  = col_double(),
    presence_binary      = col_logical(),
    needs_review         = col_logical(),
    .default             = col_character()
  )
)

# ── Cleaning ──────────────────────────────────────────────────────────────────
# 1. Drop raw_value (identical to domin_value, confirmed redundant)
# 2. Add constancy_class_numeric (I–V → 1–5); handle pipe-separated values
# 3. Create geo_releve_id: unique plot ID ordered SW→NE (smallest lat + most
#    negative lon = 1), as requested by PI

survey <- survey_raw %>%
  select(-any_of("raw_value")) %>%
  mutate(
    # Primary constancy class (take first value when pipe-separated e.g. "III|I")
    constancy_primary         = str_split_fixed(constancy_class, "\\|", 2)[, 1],
    constancy_class_numeric   = c("I"=1,"II"=2,"III"=3,"IV"=4,"V"=5)[constancy_primary],
    constancy_class_numeric_2 = {
      sec <- str_split_fixed(constancy_class, "\\|", 2)[, 2]
      ifelse(sec == "", NA_real_, c("I"=1,"II"=2,"III"=3,"IV"=4,"V"=5)[sec])
    }
  ) %>%
  select(-constancy_primary)

# Build geo_releve_id lookup
releve_uid_map <- survey %>%
  distinct(latitude, longitude, table_id, releve_id) %>%
  arrange(latitude, longitude) %>%          # SW → NE
  mutate(geo_releve_id = row_number())

survey <- survey %>%
  left_join(releve_uid_map,
            by = c("latitude", "longitude", "table_id", "releve_id"))
```

```{r overview-table}
# Counts for the "at a glance" table
releves_coords <- survey %>%
  filter(!is.na(latitude), !is.na(longitude)) %>%
  distinct(geo_releve_id) %>% nrow()

tibble(
  Metric = c("Observation rows", "Survey plots (releves)", "Vegetation classes",
             "Alliances", "Associations", "Species (standardised)",
             "Releves with coordinates"),
  Value  = c(
    format(nrow(survey), big.mark = ","),
    n_distinct(survey$geo_releve_id),
    n_distinct(survey$class,       na.rm = TRUE),
    n_distinct(survey$alliance,    na.rm = TRUE),
    n_distinct(survey$association, na.rm = TRUE),
    n_distinct(survey$species,     na.rm = TRUE),
    releves_coords
  )
) %>%
  kable(caption = "Table 1. Dataset at a glance.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

Each releve is assigned a geographic index (`geo_releve_id`) from
**1** (south-westernmost) to **`r n_distinct(survey$geo_releve_id)`**
(north-easternmost), sorting on latitude then longitude to follow
Skye's SW→NE orientation.

---


# 3 · Vegetation Classes & Associations

```{r veg-classes}
# Count unique releves per vegetation class
class_counts <- survey %>%
  distinct(geo_releve_id, class) %>%
  count(class, name = "n_releves") %>%
  arrange(desc(n_releves))

p_class <- ggplot(class_counts,
       aes(x = reorder(class, n_releves), y = n_releves, fill = n_releves)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "D") +
  labs(title = "Releves per Vegetation Class", x = NULL, y = "Number of releves") +
  theme_minimal(base_size = 11)

ggplotly(p_class)
```
```{r top-assoc}
top_assoc <- survey %>%
  filter(!is.na(association)) %>%
  distinct(geo_releve_id, association, class) %>%
  count(association, class, name = "n_releves") %>%
  slice_max(n_releves, n = 13, with_ties = FALSE)   # 13 to match Bhagyesh

p_assoc <- ggplot(top_assoc,
       aes(x = reorder(association, n_releves), y = n_releves, fill = class)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 10 Associations by Releve Count",
       x = NULL, y = "Number of releves", fill = "Class") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7))

ggplotly(p_assoc)
```

```{r full-assoc-table}
survey %>%
  filter(!is.na(association)) %>%
  distinct(geo_releve_id, class, alliance, association) %>%
  count(class, alliance, association, name = "n_releves") %>%
  arrange(desc(n_releves)) %>%
  mutate(` ` = row_number()) %>%
  select(` `, everything()) %>%
  datatable(caption  = "Table 2. Releve count by association (all associations).",
            filter   = "top",
            rownames = FALSE,
            options  = list(pageLength = 15, scrollX = TRUE))
```
---

# 4 · Species Frequency

```{r top-species}
# Count how many releves each species appears in (presence only)
species_freq <- survey %>%
  filter(presence_binary == TRUE) %>%
  count(species, name = "n_releves") %>%
  arrange(desc(n_releves))

p_top <- species_freq %>%
  slice_max(n_releves, n = 10) %>%
  ggplot(aes(x = reorder(species, n_releves), y = n_releves, fill = n_releves)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "B") +
  labs(title = "Top 10 Most-Frequent Species", x = NULL, y = "Number of releves") +
  theme_minimal(base_size = 12)

ggplotly(p_top)
```

```{r species-summary-table}
survey %>%
  filter(presence_binary == TRUE, !is.na(domin_value)) %>%
  group_by(species) %>%
  summarise(
    releves_n      = n_distinct(geo_releve_id),
    associations_n = n_distinct(association, na.rm = TRUE),
    mean_domin     = round(mean(domin_value, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  arrange(desc(releves_n)) %>%
  slice_head(n = 10) %>%
  mutate(rank = row_number()) %>%
  select(rank, everything()) %>%
  kable(col.names = c(" ", "Species", "Releves (n)", "Associations (n)", "Mean Domin"),
        caption   = "Table 3. Top 10 species by releve occurrence.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

```{r rare-species}
p_rare <- species_freq %>%
  filter(n_releves == 1) %>%          # truly rare: seen in only 1 releve
  slice_head(n = 10) %>%
  ggplot(aes(x = reorder(species, n_releves), y = n_releves, fill = n_releves)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "A") +
  labs(title = "10 Rarest Species (Globally)", x = NULL, y = "Number of releves") +
  theme_minimal(base_size = 12)

ggplotly(p_rare)
```
---

# 5 · Top-10 & Rarest-10 Species per Association

Each table shows the top 10 (or rarest 10) species within each
association, ranked by % of that association's releves.

```{r per-assoc-freq}
# For each association: how many releves contain each species?
per_assoc <- survey %>%
  filter(!is.na(association), presence_binary == TRUE) %>%
  group_by(association) %>%
  mutate(n_releves_assoc = n_distinct(geo_releve_id)) %>%
  group_by(association, species, n_releves_assoc) %>%
  summarise(n = n_distinct(geo_releve_id), .groups = "drop") %>%
  mutate(pct = round(100 * n / n_releves_assoc, 1))
```

```{r top10-per-assoc}
per_assoc %>%
  group_by(association) %>%
  slice_max(pct, n = 10, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(association, desc(pct)) %>%
  mutate(` ` = row_number()) %>%
  select(` `, Class = association, Association = association, 
         species, n_releves = n,
         assoc_releves = n_releves_assoc, pct_present = pct) %>%
  datatable(caption  = "Table 4. Top-10 species per association.",
            filter   = "top",
            rownames = FALSE,
            options  = list(pageLength = 20, scrollX = TRUE))
```


```{r rare10-per-assoc}
per_assoc %>%
  group_by(association) %>%
  slice_min(pct, n = 10, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(association, pct) %>%
  mutate(` ` = row_number()) %>%
  select(` `, association, species, n_releves = n,
         assoc_releves = n_releves_assoc, pct_present = pct) %>%
  datatable(caption  = "Table 5. 10 rarest species per association.",
            filter   = "top",
            rownames = FALSE,
            options  = list(pageLength = 20, scrollX = TRUE))
```

---

# 6 · Domin Cover Scores

The Domin scale runs 1 → 10: from sparse presence (<4% cover) to
near-complete dominance (96–100%).
```{r domin-dist}
domin_labels <- c(
  "1" = "1 (<4%)",
  "2" = "2 (4–10%)",
  "3" = "3 (11–25%)",
  "4" = "4 (26–33%)",
  "5" = "5 (34–40%)",
  "6" = "6 (41–50%)",
  "7" = "7 (51–75%)",
  "8" = "8 (76–90%)",
  "9" = "9 (91–95%)",
  "10" = "10 (96–100%)"
)

p_domin <- survey %>%
  filter(!is.na(domin_value), domin_value > 0) %>%
  mutate(domin_cat = factor(
    domin_labels[as.character(as.integer(domin_value))],
    levels = domin_labels
  )) %>%
  count(domin_cat) %>%
  ggplot(aes(x = domin_cat, y = n, fill = domin_cat)) +
  geom_col(show.legend = FALSE) +
  scale_fill_brewer(palette = "Greens") +
  labs(title = "Domin cover score distribution, all species and releves",
       x = "Domin score", y = "Observations") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

ggplotly(p_domin)
```
```{r domin-by-assoc, fig.height=10}
survey %>%
  filter(association %in% top_assoc$association,
         !is.na(domin_value), domin_value > 0) %>%
  mutate(domin_cat = factor(
    domin_labels[as.character(as.integer(domin_value))],
    levels = domin_labels
  )) %>%
  count(association, domin_cat) %>%
  ggplot(aes(x = domin_cat, y = n, fill = domin_cat)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~association, scales = "free_y", ncol = 2) +
  scale_fill_brewer(palette = "Greens") +
  labs(title = "Domin Distribution — Top 10 Associations",
       x = "Domin score", y = "Observations") +
  theme_minimal(base_size = 9) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1, size = 7))
```

```{r domin-mean-species}
# Which species have the highest mean Domin score? (min 10 occurrences)
survey %>%
  filter(!is.na(domin_value), domin_value > 0) %>%
  group_by(species) %>%
  summarise(mean_domin = mean(domin_value, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  filter(n >= 10) %>%
  slice_max(mean_domin, n = 10) %>%
  ggplot(aes(x = reorder(species, mean_domin), y = mean_domin, fill = mean_domin)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_viridis_c(option = "C") +
  labs(title = "Top 10 Species by Mean Domin Score (min 10 occurrences)",
       x = NULL, y = "Mean Domin score") +
  theme_minimal(base_size = 12)
```

---

# 7 · Community Matrix & Clustering

Binary presence/absence matrix (species in ≥ 3 releves only),
clustered with Ward's D2 on Jaccard dissimilarities.

```{r community-matrix}
# Build wide presence/absence matrix: rows = releves, cols = species
comm_long <- survey %>%
  filter(presence_binary == TRUE, !is.na(geo_releve_id)) %>%
  distinct(geo_releve_id, species)

# Keep only species found in 3 or more releves
keep_spp <- comm_long %>%
  count(species) %>%
  filter(n >= 3) %>%
  pull(species)

comm_wide <- comm_long %>%
  filter(species %in% keep_spp) %>%
  mutate(present = 1L) %>%
  pivot_wider(names_from  = species,
              values_from = present,
              values_fill = 0L) %>%
  column_to_rownames("geo_releve_id")

tibble(
  Metric  = c("Releves (rows)", "Species retained (cols)",
              "Species excluded (< 3 releves)", "Matrix fill %"),
  Value   = c(
    nrow(comm_wide),
    ncol(comm_wide),
    n_distinct(comm_long$species) - length(keep_spp),
    round(100 * sum(comm_wide) / prod(dim(comm_wide)), 1)
  )
) %>%
  kable(caption = "Table 6. Community matrix dimensions.") %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = FALSE)
```

```{r clustering, fig.height=8}
set.seed(42)
jac_dist <- vegdist(comm_wide, method = "jaccard", binary = TRUE)
hc        <- hclust(jac_dist, method = "ward.D2")
K         <- 10
clusters  <- cutree(hc, k = K)
cluster_cols <- setNames(
  colorRampPalette(brewer.pal(10, "Paired"))(K),
  as.character(1:K)
)

par(mar = c(2, 3, 4, 1))
plot(as.dendrogram(hc), leaflab = "none",
     main = glue("Hierarchical clustering (Ward / Jaccard) . k = {K} clusters"),
     ylab = "Height (Jaccard dissimilarity)")
rect.hclust(hc, k = K, border = cluster_cols)
legend("topright", legend = paste("Cluster", 1:K),
       fill = cluster_cols, cex = 0.75, ncol = 2, bty = "n")
```

```{r cluster-assoc-table}
cluster_df <- tibble(
  geo_releve_id = as.integer(names(clusters)),
  cluster       = clusters
)

# Also create sp_mat alias so section 9 works
sp_mat <- comm_wide

survey %>%
  distinct(geo_releve_id, association, class) %>%
  inner_join(cluster_df, by = "geo_releve_id") %>%
  filter(!is.na(association)) %>%
  count(Class = class, Association = association, cluster) %>%
  pivot_wider(names_from = cluster, values_from = n,
              values_fill = 0L, names_prefix = "C") %>%
  arrange(Class, Association) %>%
  datatable(
    caption = glue("Table 7. Releve count per association x cluster (k = {K})."),
    options = list(pageLength = 20),
    class   = "stripe hover compact"
  )
```

---

# 7 · Cluster Ecological Profiles

<div class="section-note">
This section characterises the `r K` Ward/Jaccard clusters from Section 6.
</div>

## 7.1 · Are the Clusters Statistically Meaningful?

```{r cluster-permanova}
set.seed(42)
cluster_fac   <- factor(clusters[rownames(sp_mat)])
permanova_res <- adonis2(jac_dist ~ cluster_fac, permutations = 199)

tibble(
  Term      = c("Cluster membership", "Residual", "Total"),
  Df        = permanova_res$Df,
  `Sum Sq`  = round(permanova_res$SumOfSqs, 2),
  R2        = round(permanova_res$R2, 3),
  F         = round(permanova_res$F, 2),
  `p-value` = permanova_res$`Pr(>F)`
) %>%
  kable(caption = glue("Table 7a. PERMANOVA (199 permutations, k = {K})."),
        digits = 3, na.rm = TRUE) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
```



```{r indicspecies, message=FALSE, warning=FALSE}
have_indic <- requireNamespace("indicspecies", quietly = TRUE)
if (have_indic) {
  set.seed(42)
  indic_res <- indicspecies::multipatt(
    sp_mat, cluster_fac, func = "r.g", duleg = TRUE,
    control = permute::how(nperm = 199)
  )
  indic_tbl <- indic_res$sign %>%
    as_tibble(rownames = "species") %>%
    filter(p.value <= 0.05, !is.na(p.value)) %>%
    mutate(cluster = as.integer(index)) %>%
    select(cluster, species, stat, p.value) %>%
    arrange(cluster, desc(stat))
} else {
  indic_tbl <- tibble(cluster = integer(), species = character(),
                       stat = double(), p.value = double())
}
```

```{r indicspecies-table}
if (nrow(indic_tbl) > 0) {
  indic_tbl %>%
    group_by(cluster) %>%
    slice_max(stat, n = 5, with_ties = FALSE) %>%
    ungroup() %>%
    rename(Cluster = cluster, Species = species,
           `Indicator value` = stat, `p-value` = p.value) %>%
    mutate(`Indicator value` = round(`Indicator value`, 3),
           `p-value` = round(`p-value`, 3)) %>%
    datatable(
      caption = "Table 7b. Statistically significant indicator species per cluster (p <= 0.05, top 5 per cluster).",
      filter  = "top",
      options = list(pageLength = 15),
      class   = "stripe hover compact"
    )
} else {
  cat("No indicator species reached p <= 0.05.")
}
```

## 7.2 · Diversity and Spread by Cluster

```{r cluster-diversity}
div_df <- tibble(
  geo_releve_id = as.integer(rownames(sp_mat)),
  richness      = specnumber(sp_mat),
  shannon       = diversity(sp_mat, index = "shannon"),
  simpson       = diversity(sp_mat, index = "simpson")
) %>%
  mutate(evenness = if_else(richness > 1, shannon / log(richness), NA_real_)) %>%
  left_join(cluster_df, by = "geo_releve_id") %>%
  left_join(survey %>% distinct(geo_releve_id, class, association),
            by = "geo_releve_id")

cluster_summary <- div_df %>%
  group_by(cluster) %>%
  summarise(
    n_releves      = n(),
    mean_richness  = round(mean(richness), 1),
    sd_richness    = round(sd(richness), 1),
    mean_shannon   = round(mean(shannon), 2),
    mean_simpson   = round(mean(simpson), 2),
    mean_evenness  = round(mean(evenness, na.rm = TRUE), 2),
    dominant_class = names(sort(table(class), decreasing = TRUE))[1],
    dominant_assoc = names(sort(table(association), decreasing = TRUE))[1],
    .groups = "drop"
  ) %>%
  arrange(desc(n_releves))
```


```{r cluster-diversity-spread, fig.height=6}
p_div_box <- ggplot(div_df, aes(x = factor(cluster), y = richness,
                                 fill = factor(cluster))) +
  geom_boxplot(show.legend = FALSE, outlier.size = 1, outlier.alpha = 0.5,
               colour = "#555555") +
  scale_fill_manual(values = cluster_cols) +
  labs(title = "Species richness per releve, by cluster",
       x = "Cluster", y = "Species richness") +
  theme_skye(10)

p_div_shannon <- ggplot(div_df, aes(x = factor(cluster), y = shannon,
                                      fill = factor(cluster))) +
  geom_boxplot(show.legend = FALSE, outlier.size = 1, outlier.alpha = 0.5,
               colour = "#555555") +
  scale_fill_manual(values = cluster_cols) +
  labs(title = "Shannon diversity per releve, by cluster",
       x = "Cluster", y = "Shannon diversity (H')") +
  theme_skye(10)

p_div_box / p_div_shannon
```

### Per-Cluster Summary Table

```{r cluster-summary-table}
cluster_summary %>%
  rename(Cluster = cluster, `Releves (n)` = n_releves,
         `Mean richness` = mean_richness, `SD richness` = sd_richness,
         `Mean Shannon` = mean_shannon, `Mean Simpson` = mean_simpson,
         `Mean evenness` = mean_evenness,
         `Dominant class` = dominant_class,
         `Dominant association` = dominant_assoc) %>%
  datatable(
    caption = "Table 7c. Diversity, spread, and dominant vegetation unit per cluster.",
    options = list(pageLength = 10, dom = "t"),
    class   = "stripe hover compact"
  )
```

### Narrative Summary - Top 3 Clusters by Size
```{r cluster-narrative, results='asis'}
top3_clusters <- cluster_summary %>% slice_head(n = 3) %>% pull(cluster)

for (cl in top3_clusters) {
  cl_row <- cluster_summary %>% filter(cluster == cl)
  top_sp <- div_df %>%
    filter(cluster == cl) %>%
    pull(geo_releve_id) %>%
    as.character() %>%
    { sp_mat[., , drop = FALSE] } %>%
    colSums() %>%
    sort(decreasing = TRUE) %>%
    head(5)
  top_sp_str <- paste(glue("*{names(top_sp)}* ({round(100 * top_sp / cl_row$n_releves, 0)}%)"),
                      collapse = ", ")
  indic_str <- indic_tbl %>%
    filter(cluster == cl) %>%
    slice_max(stat, n = 3, with_ties = FALSE) %>%
    pull(species)
  indic_str <- if (length(indic_str) > 0) paste(glue("*{indic_str}*"), collapse = ", ") else "none reached p <= 0.05"

  cat(glue(
    "\n\n**Cluster {cl}** ({cl_row$n_releves} releves, dominated by *{cl_row$dominant_class}* -- mostly the *{cl_row$dominant_assoc}* association): ",
    "mean species richness {cl_row$mean_richness} (Shannon H' = {cl_row$mean_shannon}). ",
    "Most frequent species: {top_sp_str}. ",
    "Statistically significant indicator species: {indic_str}.\n"
  ))
}
```

## 7.3 · What Dominates Each Cluster

```{r cluster-dominant-species, fig.height=20, fig.width=12}
cluster_top_sp <- div_df %>%
  select(geo_releve_id, cluster) %>%
  left_join(
    survey %>% filter(presence_binary == TRUE) %>%
      select(geo_releve_id, species),
    by = "geo_releve_id"
  ) %>%
  count(cluster, species, name = "n") %>%
  left_join(cluster_summary %>% select(cluster, n_releves), by = "cluster") %>%
  mutate(pct = round(100 * n / n_releves, 0)) %>%
  group_by(cluster) %>%
  slice_max(pct, n = 8, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(sp_label = paste(cluster, species, sep = "___"),
         sp_label = fct_reorder(sp_label, pct))

ggplot(cluster_top_sp,
       aes(x = pct, y = sp_label, fill = factor(cluster))) +
  geom_col(width = 0.75, show.legend = FALSE) +
  scale_fill_manual(values = cluster_cols) +
  scale_y_discrete(labels = function(x) str_remove(x, ".*___")) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25))) +
  facet_wrap(~ glue("Cluster {cluster}"), scales = "free_y", ncol = 2) +
  labs(title = glue("Dominant species per cluster (k = {K})"),
       x = "% of releves in cluster", y = NULL) +
  theme_skye(9) +
  theme(strip.text  = element_text(size = 9, face = "bold"),
        axis.text.y = element_text(size = 7, face = "italic"))
```

## 7.4 · Cluster Composition - Stacked Proportions

```{r cluster-stacked-class, fig.height=6}
stacked_class_df <- div_df %>%
  count(cluster, class, name = "n") %>%
  group_by(cluster) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

class_pal_local <- colorRampPalette(brewer.pal(9, "Paired"))(n_distinct(stacked_class_df$class))

ggplot(stacked_class_df,
       aes(x = factor(cluster), y = prop, fill = class)) +
  geom_col(position = "fill", width = 0.8) +
  scale_fill_manual(values = class_pal_local, name = "Vegetation class") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Vegetation-class composition by cluster",
       x = "Cluster", y = "Proportion of releves") +
  theme_skye(10) +
  theme(legend.text = element_text(size = 7), legend.position = "right")
```


# 10 · Releve Reference Table


Full geographic index of all releves.. Filter, search, or sort any column.


```{r releve-table}
survey %>%
  distinct(geo_releve_id, ref_code, table_id, latitude, longitude,
           altitude_m, slope_deg, aspect_deg, class, association) %>%
  arrange(geo_releve_id) %>%
  datatable(
    caption = "Table 8. Complete releve index, SW to NE.",
    filter  = "top",
    options = list(pageLength = 25, scrollX = TRUE),
    class   = "stripe hover compact"
  )
```
---

*Analysis by Bhagyesh Sagole & Pawan Kumar · Data: Birks H.J.B. (1973)*
